---
title: "Final project"
author: "Sille Hasselbalch Markussen"
date: "11/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Loading libraries and data}
pacman::p_load(tidyverse, lubridate)
```

## Ravelry API
In order to scrape data from Ravelry one needs a Ravelry developer account.
The associated username and password must be set in an .renviron file. 
Finally the ravelRy package must be installed and loaded.
```{r}
#install.packages('ravelRy')
library(ravelRy)
```

Get more info from this GitHub: https://github.com/walkerkq/ravelRy

## Example data acquisition, cleaning and visualisation
### DENMARK

```{r Scraping all serach results of advanced pattern search}
dk <- search_patterns(craft = "knitting", designer_country = "denmark", availability = "online", page_size = 1649, pattern_source_type = "website")
```

```{r top 10 designers}
dk_des_10 <- dk %>% 
  group_by(designer.name) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  slice(1:10)

dk_top10 <- dk %>% 
  filter(designer.name %in% dk_des_10$designer.name)
```

```{r getting info for top 10 designers}
dk_10 <- get_patterns(dk_top10$id)
dk_10$designer <- dk_top10$designer.name
```

```{r extracting years}
dk_10$published_year_short <- format(as.Date(dk_10$published, format="%Y/%m/%d"),"%y")
dk_10$published_year_long <- format(as.Date(dk_10$published, format="%Y/%m/%d"),"%Y")
```

```{r bar plot}
ggplot(dk_10, aes(published_year_short))+
  geom_bar()
```

## Data acqusition, cleaning and visualisation for remaining countries 
8 European countries were chosen with varying population size and geographical location to somewhat reflect the variety of countries in Europe. og så skulle det være lande med ikke kun 3 søgeresultater og så 10 fordi også dk
1-5: France
6-10: Spain
11-15: Czech Republic
16-20: Austria
21-25: Norway 
26-30: Ireland
31-35: Latvia
36-40: Iceland
And then my home country Denmark.

```{r}
fr <- search_patterns(craft = "knitting", designer_country = "france", availability = "online", page_size = 1450, pattern_source_type = "website")

es <- search_patterns(craft = "knitting", designer_country = "spain", availability = "online", page_size = 571, pattern_source_type = "website")

cz <- search_patterns(craft = "knitting", designer_country = "czech-republic", availability = "online", page_size = 68, pattern_source_type = "website")

at <- search_patterns(craft = "knitting", designer_country = "austria", availability = "online", page_size = 166, pattern_source_type = "website")

no <- search_patterns(craft = "knitting", designer_country = "norway", availability = "online", page_size = 1521, pattern_source_type = "website")

ie <- search_patterns(craft = "knitting", designer_country = "ireland", availability = "online", page_size = 479, pattern_source_type = "website")

lv <- search_patterns(craft = "knitting", designer_country = "latvia", availability = "online", page_size = 98, pattern_source_type = "website")

is <- search_patterns(craft = "knitting", designer_country = "iceland", availability = "online", page_size = 165, pattern_source_type = "website")

```

```{r function for top 10 designers}
top_10 <- function(country){
  designer10 <- country %>% 
  group_by(designer.name) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  slice(1:10)

  country10 <- country %>% 
    filter(designer.name %in% designer10$designer.name)
}
```

```{r top 10 for each country in countries}
fr10 <- top_10(fr)
es10 <- top_10(es)
cz10 <- top_10(cz)
at10 <- top_10(at)
no10 <- top_10(no)
ie10 <- top_10(ie)
lv10 <- top_10(lv)
is10 <- top_10(is)
```

```{r extracting pattern info}
fr_10 <- get_patterns(fr10$id)
fr_10$designer <- fr10$designer.name

es_10 <- get_patterns(es10$id)
es_10$designer <- es10$designer.name

cz_10 <- get_patterns(cz10$id)
cz_10$designer <- cz10$designer.name

at_10 <- get_patterns(at10$id)
at_10$designer <- at10$designer.name

no_10 <- get_patterns(no10$id)
no_10$designer <- no10$designer.name

ie_10 <- get_patterns(ie10$id)
ie_10$designer <- ie10$designer.name

lv_10 <- get_patterns(lv10$id)
lv_10$designer <- lv10$designer.name

is_10 <- get_patterns(is10$id)
is_10$designer <- is10$designer.name
```

```{r column with country}
fr_10$country <- "France"
es_10$country <- "Spain"
cz_10$country <- "Czech Republic"
at_10$country <- "Austria"
no_10$country <- "Norway"
ie_10$country <- "Ireland"
lv_10$country <- "Latvia"
is_10$country <- "Iceland"
dk_10$country <- "Denmark"
```

```{r combining to one df}
countries <- rbind(fr_10, es_10, cz_10, at_10, no_10, ie_10, lv_10, is_10)
```

```{r extracting years}
countries$published_year_short <- format(as.Date(countries$published, format="%Y/%m/%d"),"%y")
countries$published_year_long <- format(as.Date(countries$published, format="%Y/%m/%d"),"%Y")
```

```{r adding Denmark}
# Since DK was used as an example for the method and already has a published year column it is added to the df here
countries <- rbind(countries, dk_10)
```

```{r getting rid of NAs}
countries <- countries %>% drop_na(published_year_long)
```

```{r writing csv}
# The extract pattern info is quite heavy to run (takes > 3 hours on my MacBook 12 inch)
countries <- countries %>% 
  select(published, published_year_long, published_year_short, country, designer, name, product_id)

write_csv(countries, "ravelry_knit_data_clean.csv")
```


## Lockdown dates per country
```{r read data}
lockdown_dates <- read_csv("countryLockdowndates.csv")
```

```{r filtering countries}
country_list <- c("France", "Spain", "Czechia", "Austria", "Norway", "Ireland", "Latvia", "Iceland", "Denmark")
lockdown_dat <- lockdown_dates %>% 
  filter(`Country/Region` %in% country_list)
```

```{r filtering provinces}
# All rows that has NA in the Province column indicates a countrywide lockdown 
lockdown_dat <- lockdown_dat %>% 
  filter(is.na(Province))
```

```{r changing date format}
lockdown_dat$Date <- as.Date(lockdown_dat$Date, format = "%d/%m/%Y")
```


```{r writing csv}
write_csv(lockdown_dat, "lockdown_data_clean.csv")
```

